---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Analaysing the Persistence and Patterns of Baby Naming Conventions"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Nico Katzke}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Annika Pretorius^[__Contributions:__  \\newline _The authors would like to thank the World Happiness survey for showing us what happiness is._]"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch" # First Author's Affiliation
Email1: "21626979@sun.ac.za" # First Author's Email address

CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  This article aims to discuss what makes us happy. This is a suggested solution to question 1.
---


```{r setup, include=FALSE}

library(tidyverse)
library(stringr)
list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

#Loading the Data and Creating Different Datasets for Later Use
Baby_Names <- readRDS("Data/US_Baby_names/Baby_Names_By_US_State.rds")
charts <- readRDS("Data/US_Baby_names/charts.rds")
HBO_Titles <- readRDS("Data/US_Baby_names/HBO_titles.rds")
HBO_Credits <- readRDS("Data/US_Baby_names/HBO_credits.rds")

Baby_Names_df<- filter_and_arrange_baby_names(Baby_Names)
HBO_joined_df <- prepare_HBO_data(HBO_Titles, HBO_Credits)
combined_result_final_top_25_names <- combine_top_25_names(Baby_Names_df, HBO_joined_df, "Count", "tmdb_popularity")

Girl_Baby_Names_df <- Baby_Names_df %>% filter(Gender == "F")
Boy_Baby_Names_df <- Baby_Names_df %>% filter(Gender == "M")


```

Firstly, lets analyse baby names nationally, rnak the top 25 per year and compare it to the HBO titles from 1950 to 2014 

```{r,  echo = FALSE, warning =  TRUE, fig.align = 'center', fig.cap = "Total Count of Baby Names by Gender.\\label{Figure1}", fig.ext = 'png', fig.height = 5, fig.width = 7}

g <- plot_total_baby_names_per_year_by_gender(Girl_Baby_Names_df, Boy_Baby_Names_df)
print(g)

```


```{r,  echo = FALSE, warning =  TRUE, fig.align = 'center', fig.cap = "The Proportion of the Top 25 Baby Names and their Persistence Over Time\\label{Figure1}", fig.ext = 'png', fig.height = 5, fig.width = 7}
#The vertical axis represents the proportion of babies given a specific name each year. This is calculated as the count of babies with that name divided by the total count of babies for that year.

g1<- plot_top_25_name_proportions(Girl_Baby_Names_df, Boy_Baby_Names_df)
g1
```
The figure above illustrates the proportion of the top 25 baby names over time for both female and male names.The horizontal axis represents the years from 1950 to 2014. The vertical axis represents the proportion of babies given a specific name each year. This is calculated as the count of babies with that name divided by the total count of babies for that year.Each line represents a specific name that was in the top 25 for that year. The color of each line corresponds to a different name. The plot is divided into two facets: one for female names and one for male names. This separation helps to compare the trends for female and male names independently.

+ Female Names: The left facet shows the proportion trends for the top 25 female names over time. The plot indicates that the popularity of the top names was relatively higher in the past (1950s to 1980s) and has become more evenly distributed in recent years.

+ Male Names: The right facet shows the proportion trends for the top 25 male names over time. Similar to female names, the plot indicates that the popularity of the top names was higher in the past and has become more evenly distributed in recent years. 

The peaks in the plot indicate years when specific names were particularly popular. The valleys indicate years when the popularity of those names decreased. Declining Trends: For both female and male names, the plot shows a general declining trend in the proportion values, indicating that the popularity of top names has become less concentrated over time. This suggests a diversification in naming trends, with a wider variety of names being used in recent years.

#Analysing the Time Series Plot


```{r}
library(dplyr)
library(tidyverse)
library(stringr)

# Find common names between Baby Names and HBO Titles
common_names <- combined_result_final %>%
    filter(!is.na(Name_BabyNames) & !is.na(Name_HBO) & Name_BabyNames == Name_HBO) %>%
    group_by(Year) %>%
    summarize(CommonNamesCount = n())


rank_correlation <- combined_result_final %>%
    group_by(Year) %>%
    summarise(Correlation = cor(Top_25_Girl_names, Name_HBO, method = "spearman"))


```


```{r}
Top_25_Girl_names <- Top_25_Girl_names %>%
  mutate(Gender = "F")

Top_25_Boy_names <- Top_25_Boy_names %>%
  mutate(Gender = "M")

# Combine the data frames
Top_25_names <- bind_rows(Top_25_Girl_names, Top_25_Boy_names)

# Create a data frame indicating presence in top 25
name_persistence <- Top_25_names %>%
  mutate(Presence = 1) %>%  # Add a column to indicate presence
  select(Year, Name, Gender, Presence) %>%
  pivot_wider(names_from = Year, values_from = Presence, values_fill = 0) %>%
  pivot_longer(cols = -c(Name, Gender), names_to = "Year", values_to = "Presence") %>%
  mutate(Year = as.integer(Year))

# Identify the most persistent names
persistent_names <- name_persistence %>%
  group_by(Name, Gender) %>%
  summarize(TotalPresence = sum(Presence), .groups = 'drop') %>%
  arrange(desc(TotalPresence)) %>%
  group_by(Gender) %>%
  top_n(10, TotalPresence) %>%
  pull(Name)

# Filter the name_persistence dataframe to include only the most persistent names
filtered_name_persistence <- name_persistence %>%
  filter(Name %in% persistent_names)

# Heatmap plot to show persistence of the most persistent names
g1 <- filtered_name_persistence %>% ggplot(aes(x = Year, y = Name, fill = Presence)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "blue") +
  facet_wrap(~ Gender, scales = "free_y") +
  labs(title = "Persistence of Top Baby Names Over Time",
       x = "Year",
       y = "Name",
       fill = "Presence") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8))  # Adjust y-axis text size for readability

g1

```


Summarizing how long each name stayed in top 25 
```{r}
# Summarize the number of years each name stayed in the top 25
name_persistence_summary <- name_persistence %>%
  group_by(Name, Gender) %>%
  summarize(BabyNamesYearsInTop25 = sum(Presence), .groups = 'drop') %>%
  arrange(desc(BabyNamesYearsInTop25))

# Filter the name_persistence_summary dataframe to include only the most persistent names
filtered_summary <- name_persistence_summary %>%
  group_by(Gender) %>%
  top_n(10, BabyNamesYearsInTop25)


# Bar plot to show the number of years each name stayed in the top 25
g2 <- filtered_summary %>% ggplot(aes(x = reorder(Name, BabyNamesYearsInTop25), y = BabyNamesYearsInTop25, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  labs(title = "Number of Years in Top 25 Baby Names",
       x = "Name",
       y = "Years in Top 25",
       fill = "Gender") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8)) 
g2
```

We can therefore see that the top boy names over the last few decades illustrate significant persistence in comparison to that of girl names, therefore, we will further analyse factors influencing girl names as they display significantly higher variability which could be linked to alternate factors. Additionally, this is found in figure 2 whereby we see the persistence for girls is roughly a similar number of years except for Elizabeth. 

```{r}

# Create a data frame indicating presence in top 25
name_persistence_HBO <- Top_25_HBO_names %>%
  mutate(Presence = 1) %>%
  select(Year, Name, Presence) %>%
  pivot_wider(names_from = Year, values_from = Presence, values_fill = 0) %>%
  pivot_longer(cols = -Name, names_to = "Year", values_to = "Presence") %>%
  mutate(Year = as.integer(Year))

# Identify the most persistent names
persistent_names_HBO <- name_persistence_HBO %>%
  group_by(Name) %>%
  summarize(TotalPresence = sum(Presence), .groups = 'drop') %>%
  arrange(desc(TotalPresence)) %>%
  top_n(10, TotalPresence) %>%
  pull(Name)

# Filter the name_persistence_HBO dataframe to include only the most persistent names
filtered_name_persistence_HBO <- name_persistence_HBO %>%
  filter(Name %in% persistent_names_HBO)

# Heatmap plot to show persistence of the most persistent names
g3 <- filtered_name_persistence_HBO %>% 
  ggplot(aes(x = Year, y = Name, fill = Presence)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = "Persistence of Top HBO Names Over Time",
       x = "Year",
       y = "Name",
       fill = "Presence") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8))

# Display the plot
g3

```
We can see the above persistence is all male names 

```{r}
HBO_name_persistence_summary <- name_persistence_HBO %>%
  group_by(Name) %>%
  summarize(HBONamesYearsInTop25 = sum(Presence), .groups = 'drop') %>%
  arrange(desc(HBONamesYearsInTop25))

# Filter the summary dataframe to include only the most persistent names
filtered_HBO_summary <- HBO_name_persistence_summary %>%
  top_n(10, HBONamesYearsInTop25)

# Bar plot to show the number of years each HBO name stayed in the top rankings
g4 <- filtered_HBO_summary %>% ggplot(aes(x = reorder(Name, HBONamesYearsInTop25), y = HBONamesYearsInTop25)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Number of Years HBO Names Stayed in Top Rankings",
       x = "HBO Name",
       y = "Years in Top Rankings") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8))

# Display the plot
g4

```
Now i will analyse the correlation between the persistent boy names within the Baby names dataframe in comparison to that of the most persistent names within the HBO baby frame, which all happen to be male. 

```{r}
# Rename columns to have consistent names for merging
filtered_babyname_summary <- filtered_summary %>% rename(BabyNamesYearsInTop25= YearsInTop25)

filtered_HBO_summary_final <- filtered_HBO_summary %>% rename(HBONamesYearsinTop25 = YearsInTop)
# Find common names between baby names and HBO names

common_names <- inner_join(filtered_babyname_summary, filtered_HBO_summary_final, by = "Name")

common_names2 <- inner_join(HBO_name_persistence_summary,name_persistence_summary, by = "Name")

colnames(common_names2)
```


```{r}
# Prepare data for plotting
common_names_long <- common_names %>%
  pivot_longer(cols = c(BabyNamesYearsInTop25, HBONamesYearsinTop25), 
               names_to = "Source", 
               values_to = "YearsInTop25") %>%
  mutate(Source = recode(Source, 
                         "BabyNamesYearsInTop25" = "Baby Names",
                         "HBONamesYearsInTop25" = "HBO Names"))

# Plot the side-by-side bar chart
g5 <- ggplot(common_names_long, aes(x = Name, y = YearsInTop25, fill = Source)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Persistence of Common Names in Top 25 Rankings",
       x = "Name",
       y = "Years in Top 25 Rankings",
       fill = "Source") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

g5

```



```{r}
correlation_value <- cor(common_names2$BabyNamesYearsInTop25, common_names2$HBONamesYearsInTop25, method = "spearman", use = "pairwise.complete.obs")

# Function to calculate correlation and create a plot
correlation_plot <- function(data) {
  # Calculate correlation
  cor_value <- cor(data$BabyNamesYearsInTop25, data$HBONamesYearsInTop25, method = "spearman")
  
  # Plot correlation
  ggplot(data, aes(x = BabyNamesYearsInTop25, y = HBONamesYearsInTop25)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, color = "blue") +
    labs(title = paste("Correlation between Persistence of Baby Names and HBO Names: ", round(cor_value, 2)),
         x = "Years in Top 25 Baby Names",
         y = "Years in Top HBO Rankings") +
    theme_minimal()
}

# Calculate and plot correlation
correlation_plot(common_names2)

```

```{r}
# Identify spikes in popularity for names
spikes <- combined_result_final%>%
  group_by(Name_BabyNames) %>%  
  summarise(max_count = max(Rank_BabyNames), max_year = Year[which.max(Rank_BabyNames)]) %>% arrange(max_year)

# Calculate year-over-year change in rank for baby names
spikes <- combined_result_final %>%
  group_by(Name_BabyNames) %>%
  arrange(Year) %>%
    select(Year, Name_BabyNames, Rank_BabyNames) %>% 
    mutate(YearlyChange = Rank_BabyNames - lag(Rank_BabyNames)) %>%
  filter(!is.na(YearlyChange)) %>%
  ungroup()

# Identify significant spikes (e.g., a rank improvement of more than 10)
significant_spikes <- spikes %>%
  filter(YearlyChange < -10) %>% rename(Name = Name_BabyNames) # Negative change indicates improvement in rank

# Normalize HBO title names for matching
HBO_joined_spikes <- HBO_joined %>%
  mutate(name = str_to_lower(str_trim(stri_enc_toutf8(Name))))

# Match spikes with HBO titles
influence_analysis <- significant_spikes %>%
  left_join(HBO_joined_df, by = c("Name", "Year"))

  

# Analyze influence of HBO titles and popular media
influence_analysis <- HBO_joined_df%>%
  filter(Name %in% top_25_baby_names$Name) %>% arrange(Year) %>% select(Name, Year)

```


```{r}
list.files("Data/US_Baby_namess")

# Function to check and read RDS files
read_rds_file <- function(path) {
  if (file.exists(path)) {
    readRDS(path)
  } else {
    stop(glue::glue("File {path} does not exist."))
  }
}

# Check and read each file
files <- lapply(file_paths, read_rds_file)

# Assign variables
Baby_Names <- files$Baby_Names
charts <- files$charts
HBO_Titles <- files$HBO_Titles
HBO_Credits <- files$HBO_Credits
```


