---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "The Longevity and Musical Progression of Coldplay versus Metallica"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Nico Katzke}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Annika Pretorius (21626979)^[__Contributions:__  \\newline _The authors would like to thank Spotify and the Billboard Top 100 for the data utilized within the following investigation._]"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University" # First Author's Affiliation


CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 12pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  This article aims to discuss the longevity and musical progression of two famous bands: Coldplay and Metallica
---


```{r setup, include=FALSE}

# Load necessary libraries
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(tidyverse)

list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

#Loading the Data
coldplay <- read_csv("Data/Coldplay_vs_Metallica/Coldplay.csv")
metallica <- read_csv("Data/Coldplay_vs_Metallica/metallica.csv")
spotify <- readRDS("Data/Coldplay_vs_Metallica/Broader_Spotify_Info.rds")
billboard_100 <- readRDS("Data/Coldplay_vs_Metallica/charts.rds")

#Filtering the data so only studio produced songs are present
coldplay_filtered <- filter_coldplay_data(coldplay)

metallica_filtered <- filter_metallica_data(metallica)

# Combine the datasets for comparison
combined_band_data <- bind_rows(coldplay_filtered, metallica_filtered)

```

# An Analysis of the Music Industry \label{An Analysis of the Music Industry}

Coldplay and Metallica are two iconic bands that have left an indelible mark on the music industry. Coldplay, known for its atmospheric soundscapes and introspective lyrics, has captivated audiences worldwide since its debut in the early 2000s. In contrast, Metallica's powerful riffs and intense performances have made them a staple in the heavy metal genre since the 1980s. This project aims to compare the performance of these two bands using data from the Billboard Top 100 and Spotify, providing insights into their popularity and influence over time.

The analysis focuses on several key variables within the datasets. From the Billboard Top 100, we examine metrics such as the number of songs each band has placed in the Top 100, the peak rank achieved by these songs, and the total number of weeks they have remained on the chart. These metrics offer a comprehensive view of each band's commercial success and longevity in the music charts. On Spotify, we analyze the number of streams and popularity scores of their songs, providing a modern perspective on their relevance in the current digital music landscape. By comparing these variables, we can gain a deeper understanding of how Coldplay and Metallica have evolved over time and how their music continues to resonate with audiences today.


# Comparisons of the Bands Popularity\label{Comparisons of the Bands Popularity}


Coldplay and Metallica, though distinct in their musical styles, both enjoy significant popularity across their albums. Popularity in the figures below measure the popularity of the albums ranging form 0 to 100. The Data source from which this was taken is from spotify


```{r,  echo = FALSE, warning =  FALSE, fig.align = 'center', fig.cap = "Coldplay's Popularity per Album.\\label{Figure1}", fig.ext = 'png', fig.height = 5, fig.width = 7}

g1<- popularity_boxplot(coldplay_filtered, xaxis_size = 10, xaxis_rows = 3)

g1
```

```{r,  echo = FALSE, warning =  FALSE, fig.align = 'center', fig.cap = "Metallica's Popularity per Album.\\label{Figure2}", fig.ext = 'png', fig.height = 5, fig.width = 7}

g2<- popularity_boxplot(metallica_filtered, xaxis_size = 10, xaxis_rows = 3)

g2
```
As you can see from \ref{Figure1} and \ref{Figure2}, Coldplay and Metallica, despite having very different musical styles, both enjoy massive popularity across their albums. However as seen in \ref{Figure3} below, Coldplay is more popular overall than Metallica. 

```{r,  echo = FALSE, warning =  FALSE, fig.align = 'center', fig.cap = "Comparing both Band's Popularity\\label{Figure3}", fig.ext = 'png', fig.height = 5, fig.width = 7}

g3<- create_box_plot(combined_band_data, "popularity", "Popularity Comparison: Coldplay vs Metallica")
g3

```

# Density Plot of Popularity:

The density plot reveals the distribution of song popularity for Coldplay and Metallica. Coldplay's songs tend to have a more uniform distribution with a slight peak, indicating a consistent level of popularity across their catalog. Metallica, on the other hand, shows a more pronounced peak, suggesting that a significant number of their songs reach higher popularity scores. This contrast highlights how Metallica has a core set of extremely popular tracks, while Coldplay's popularity is more evenly spread across their discography. This plot helps us understand the different ways in which each band's music resonates with their listeners.

```{r,  echo = FALSE, warning =  FALSE, fig.align = 'center', fig.cap = "Density Plot of Popularity\\label{Figure4}", fig.ext = 'png', fig.height = 5, fig.width = 7}
g5 <- combined_density_plot(combined_band_data, "popularity", "Density Plot of Popularity: Coldplay vs Metallica")
g5

```

# Scatter Plot of Energy vs Danceability:

The scatter plot of energy versus danceability, \ref{Figure5}, offers an insightful comparison between Coldplay and Metallica's musical styles. Coldplay's songs generally cluster towards the higher end of danceability, reflecting their melodic and rhythmic compositions designed to be both catchy and uplifting. In contrast, Metallica's tracks tend to score higher on energy but lower on danceability, showcasing their intense sound that's characteristic of heavy metal. This scatter plot effectively illustrates the distinct musical signatures of each band, highlighting Coldplay's tendency towards more danceable tunes and Metallica's focus on high-energy, powerful performances.

```{r,  echo = FALSE, warning =  FALSE, fig.align = 'center', fig.cap = "Energy versus Danceability\\label{Figure5}", fig.ext = 'png', fig.height = 5, fig.width = 7}
g6 <- combined_scatter_plot(combined_band_data, "energy", "danceability", "Energy vs Danceability: Coldplay vs Metallica")
g6
```


# Analysing the Billboard Top 100

Coldplay and Metallica have both spent a lot of time on the Billboard Top 100, showing their lasting popularity. Coldplay's songs, with their emotional and melodic style, have kept them on the charts for many weeks. Metallica's powerful and energetic tracks have also secured them a strong presence, especially in the rock and metal genres. The bar plot below clearly shows that while both bands have had substantial success, their chart performance underscores their ability to engage listeners and sustain their popularity over extended periods. This analysis provides a compelling visual representation of the lasting impact both bands have had on the music industry.

```{r,  echo = FALSE, warning =  FALSE, fig.align = 'center', fig.cap = "The Billboard 100s Results for both bands\\label{Figure6}", fig.ext = 'png', fig.height = 5, fig.width = 7}

# Filter the Billboard dataset for Coldplay and Metallica
billboard_filtered <- billboard_100 %>%
  filter(grepl("Coldplay", artist, ignore.case = TRUE) | grepl("Metallica", artist, ignore.case = TRUE)) %>%
  mutate(band = ifelse(grepl("Coldplay", artist, ignore.case = TRUE), "Coldplay", "Metallica")) %>% rename(weeks_on_board = `weeks-on-board`)


g7 <- bar_plot_billboard_weeks(billboard_filtered)
g7

```

\newpage

# Conclusion

In conclusion, Coldplay and Metallica have demonstrated remarkable longevity and musical progression over the years. Coldplay's ability to evolve their sound while maintaining their signature emotional and melodic style has kept them relevant and popular. Similarly, Metallica's consistent delivery of powerful, energetic music has solidified their place in the heavy metal genre. Both bands have shown that they can adapt and thrive in the ever-changing music industry, securing a lasting legacy through their continuous presence on the charts and streaming platforms.
